### 2024-11-12 paper/paper.Rmd
interactive maps added, maps section first draft

---
title: "Ride at Your Own Risk: Insights into Toronto’s Bicycle Theft Hotspots"
subtitle: "[To Be Updated....]"
author: "Yingke He, Ziheng Zhong"
thanks: "Code and data are available at: [https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main](https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main)."
date: today
output: html_document
bibliography: references.bib
---

<style>
body {
  background-color: #f0f8ff; /* AliceBlue color */
}
</style>

# introduction
write an intro to convey the issue
the goal of the article :be careful of bikes

# When Are Bikes at the Greatest Risk? A Look at Seasonal and weekly Patterns

```{r mothly_plot, echo=FALSE, warning = FALSE, message = FALSE, fig.width=11, fig.height=5}
# Load necessary libraries
library(tidyverse)
library(ggridges)
library(cowplot)
library(here)

# Load your data
data <- read_csv(here("data", "Theft_and_Found_Counts_By_Month_2014_2024.csv"), show_col_types = FALSE)

# Convert month names to month numbers for ordering
data <- data %>%
  mutate(OCC_MONTH_NUM = factor(OCC_MONTH, levels = month.name))

# Specify levels for OCC_YEAR to order from 2014 to 2024
data$OCC_YEAR <- factor(data$OCC_YEAR, levels = 2014:2024)

# Calculate a scaled height for "Found" data by dividing by the maximum "Found" value
max_found <- max(data$Found, na.rm = TRUE)
max_stolen <- max(data$Stolen, na.rm = TRUE)

# Plot for Bikes Found with scaled heights
plot_found <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                               height = (Found / max_found) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#05595B", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    plot.title = element_text(size = 10, color = "#05595B", hjust = -0.1, vjust = 1.5),  # Smaller title
    panel.background = element_rect(fill = "#f0f8ff", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#f0f8ff", color = NA)    # Light blue background for overall plot
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Found")  # Add title for Bikes Found

# Plot for Bikes Stolen with scaled heights and no borders
plot_stolen <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                                height = (Stolen / max_stolen) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#603601", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    panel.background = element_rect(fill = "#f0f8ff", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#f0f8ff", color = NA),   # Light blue background for overall plot
    plot.title = element_text(size = 10, color = "#603601", hjust = -0.1, vjust = 1.5)  # Smaller title
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Stolen")  # Add title for Bikes Stolen

# Combine the plots side by side
combined_plot <- plot_grid(plot_found, plot_stolen, nrow = 1)

# Display the combined plot
combined_plot

```

```{r interactive_weekly_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=7, fig.height=4}
library(plotly)
library(readr)
library(dplyr)
library(here)

# Load the data
data <- read_csv(here("data", "Theft_Counts_By_Day.csv"), show_col_types = FALSE)

# Arrange data by day of the week (Monday to Sunday)
data <- data %>%
  arrange(factor(OCC_DOW, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))

# Initialize plot with a line trace for the entire data
fig <- plot_ly(data, x = ~OCC_DOW, y = ~Number_of_Thefts, type = 'scatter', mode = 'lines+markers',
               line = list(color = 'brown'), marker = list(color = 'blue', size = 8))

# Add points for each day with initial visibility settings
for (i in 1:nrow(data)) {
  fig <- fig %>%
    add_trace(
      x = data$OCC_DOW[i],
      y = data$Number_of_Thefts[i],
      type = 'scatter',
      mode = 'markers+text',
      text = paste("Thefts:", data$Number_of_Thefts[i]),
      textposition = "top center",
      marker = list(color = "yellow", size = 10),
      showlegend = FALSE,
      visible = ifelse(i == 1, TRUE, FALSE)
    )
}

# Define steps for each day in the slider
steps <- lapply(1:nrow(data), function(i) {
  step <- list(
    method = "restyle",
    args = list("visible", as.list(rep(FALSE, nrow(data) + 1))),  # +1 to include the main line trace
    label = data$OCC_DOW[i]
  )
  step$args[[2]][i + 1] <- TRUE  # Make only the current day’s point visible (offset by 1 due to main line trace)
  step$args[[2]][1] <- TRUE       # Keep the main line trace visible
  return(step)
})

# Add layout with the slider and set background colors
fig <- fig %>%
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Number of Thefts"),
    sliders = list(
      list(
        active = 0,
        currentvalue = list(prefix = "Day: "),
        steps = steps
      )
    ),
    plot_bgcolor = "#f0f8ff",     # Set background color for the plotting area
    paper_bgcolor = "#f0f8ff"     # Set background color for the entire plot
  )

# Display the plot
fig

```

# What Makes a Bike a Target? Analyzing Theft by Bicycle Type and Value
Discuss any available data on bike types, brands, or values, showing if certain bikes are more vulnerable to theft than others.

# A Decade of Disappearing Bikes: Tracing Theft Trends Over Ten Years

```{r animated_bar_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=7, fig.height=4}
# Libraries
library(ggplot2)
library(gganimate)
library(readr)
library(here)
library(dplyr)

# Load the data
data <- read_csv(here("data", "Stolen_Found_Counts.csv"), show_col_types = FALSE)

# Rename "THEFT" to "Stolen" in the Offence_Category column
data <- data %>%
  mutate(Offence_Category = ifelse(Offence_Category == "THEFT", "Stolen", Offence_Category))

# Set the number of frames for the animation
max_frames <- 100  # Total number of frames for smooth animation

# Generate interpolated data for simultaneous increase
animated_data <- data %>%
  # Repeat each category for each frame
  group_by(Offence_Category) %>%
  do(data.frame(
    Offence_Category = .$Offence_Category,
    Number_of_Offences = seq(0, .$Number_of_Offences, length.out = max_frames),
    frame = 1:max_frames
  ))

# Ensure Offence_Category is a factor for proper ordering
animated_data$Offence_Category <- as.factor(animated_data$Offence_Category)

# Define custom colors for each category
custom_colors <- c("Stolen" = "#603601", "FOUND" = "#4682B4", "Other" = "#eedaaf")

# Create the animated bar plot
ggplot(animated_data, aes(x=Offence_Category, y=Number_of_Offences, fill=Offence_Category)) + 
  geom_bar(stat='identity') +
  theme_minimal() +
  labs(x = 'Offense Category', y = 'Number of Offences') +

  # Apply custom colors
  scale_fill_manual(values = custom_colors) +

  # Remove the legend title
  guides(fill = guide_legend(title = NULL)) +

  # Set background colors for panel and plot
  theme(
    panel.background = element_rect(fill = "#f0f8ff", color = NA),
    plot.background = element_rect(fill = "#f0f8ff", color = NA)
  ) +

  # gganimate specific settings for smooth transition
  transition_states(
    frame,
    transition_length = 2,
    state_length = 1
  ) +
  ease_aes('linear')  # Ensures a smooth, even increase

```
<div style="color: grey; font-size: 12 px; margin-top: 3px;"> Displays the count of various bicycle-related offences reported to the Toronto Police Service, highlighting categories such as stolen, found, and other incidents. The smooth increase in bar heights emphasizes the scale of each offence type, offering insights into the distribution and relative frequency of these incidents from 2014 to 2024 </div> \

```{r animated_line_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=6, fig.height=3}
# Load necessary libraries
library(ggplot2)
library(gganimate)
library(dplyr)
library(hrbrthemes)
library(readr)
library(here)

# Load your data
data <- read_csv(here("data", "Bicycle_Theft_Monthly_Counts.csv"), show_col_types = FALSE)

# Create a numeric month order variable
data <- data %>%
  mutate(
    OCC_MONTH_NUM = match(OCC_MONTH, c("January", "February", "March", "April", "May", 
                                       "June", "July", "August", "September", "October", 
                                       "November", "December"))
  )

# Create the animated plot
plot <- data %>%
  ggplot(aes(x = OCC_MONTH_NUM, y = Number_of_Thefts, group = as.factor(OCC_YEAR), color = as.factor(OCC_YEAR))) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d() +
  ylab("Number of Bikes Stolen") +
  xlab("Month") +
  scale_x_continuous(breaks = 1:12, labels = c("January", "February", "March", "April", "May", "June", 
                                               "July", "August", "September", "October", "November", "December")) +
  theme_ipsum() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.background = element_rect(fill = "#f0f8ff", color = NA),
    plot.background = element_rect(fill = "#f0f8ff", color = NA)
  ) +
  transition_reveal(OCC_MONTH_NUM) +
  labs(color = "Year")

# Render the animation
animate(plot, nframes = 100, fps = 10)

```
<div style="color: grey; font-size: 12 px; margin-top: 3px;"> Visualizes the monthly count of bicycle thefts reported to the Toronto Police Service from 2014 to June 2024. The animation reveals seasonal patterns and variations across different years, providing insights into peak months and potential trends over time. </div> \

# Mapping the Danger Zones: High-Risk Locations Across Toronto

Now imagine you’re navigating the streets of Toronto, trying to understand where bicycle thefts are happening. The first chart is like an overhead view from a hot air balloon, showing clusters of incidents with bright orange circles scattered across the landscape. Each circle represents a collection of theft reports—smaller bubbles indicate fewer incidents, while larger ones stand like towering beacons, signaling areas of greater vulnerability. Downtown Toronto, unsurprisingly, is covered by a massive bubble, indicating that this densely populated urban center has been most affected by bike thefts. As we drift north and away from the downtown core, the bubbles shrink, suggesting fewer incidents—likely due to the suburban nature of these areas, where bikes are stored more securely. And here's the interactive part—you can click on these bubbles to zoom in and explore specific areas, uncovering the exact locations and getting a more detailed sense of bike stolen cases. It's like having a personal guide leading you through the hot spots and quieter neighborhoods.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#### Map with Markers ####

# Load necessary libraries
library(dplyr)         # Provides functions for data manipulation.
library(leaflet)       # Enables interactive maps.
library(purrr)         # Contains functional programming tools, used here to walk through data.
library(readr)         # Provides functions to read rectangular data like CSV files.
library(here)          # Helps manage file paths in a more robust way.

# Load your dataset (adjust the file path accordingly)
map_data <- read_csv(here("data", "map_by_neighbourhoods_data.csv"), show_col_types = FALSE)
# Loads the CSV file from the specified path using the 'here()' function for consistent path management.
# 'show_col_types = FALSE' suppresses messages about the data types of each column.

# Modify the data frame, categorizing case_num column
map_data <- map_data %>% 
  mutate(
    mag.level = cut(case_num, c(100, 500, 1000, 1500),
                    labels = c('> 100 & <=500', '>500 & <=1000', '>1000 & <=1500'))
    # Creates a 'mag.level' column that categorizes 'case_num' into three levels based on specified ranges.
  ) %>%
  split(.$mag.level)  # Splits the dataset into a list of data frames by 'mag.level'.

# Initialize the leaflet map
l <- leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap)
# Initializes a Leaflet map object and adds a base tile layer from the Esri Ocean Basemap provider.

# Add the markers to the map for each case_num level
names(map_data) %>%
  walk(function(df) {
    l <<- l %>% 
      addMarkers(
        data = map_data[[df]], lng = ~longitude, lat = ~latitude,
        label = ~paste0("Hood: ", hood_name, " | Stolen Place: ", stolen_place, " | Cost: ", cost, " | Status: ", status),
        popup = ~paste0(
          "Hood: ", hood_name, "<br>",
          "Stolen Place: ", stolen_place, "<br>",
          "Cost: ", cost, "<br>",
          "Status: ", status, "<br>",
          "Case in area: ", case_num
        ),
        group = df,
        clusterOptions = markerClusterOptions()
      )
  })
# Iterates over the names of the 'map_data' list (which are the different 'mag.level' categories).
# For each category, it adds markers to the map:
# - `data = map_data[[df]]`: Uses the split data for the current case_num level.
# - `lng = ~longitude, lat = ~latitude`: Sets longitude and latitude for marker placement.
# - `label`: Displays the 'case_num' value when markers are hovered over.
# - `popup`: Displays additional information (hood_num, stolen_place, cost, status, and case_num) when a marker is clicked.
# - `group = df`: Groups the markers by 'mag.level'.
# - `clusterOptions = markerClusterOptions()`: Enables clustering of markers for better visualization.

# Add layer controls and a minimap
l <- l %>%
  addLayersControl(
    overlayGroups = names(map_data),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addMiniMap(tiles = providers$Esri.OceanBasemap, width = 120, height = 80)
# Adds layer controls to allow switching between different 'mag.level' categories on the map.
# Adds a minimap showing a small overview of the area using the Esri Ocean Basemap.

# Display the Map
l
```

In the second chart, we dive deeper into the city with a ground-level view—a sea of red dots each marks an individual theft incident. The dots are packed tightly in certain neighborhoods, sprawling across downtown Toronto like a dense red blanket. It paints a vivid picture of the density of bike thefts—evidence that the bustling central part of the city faces a higher risk. Street after street is littered with these red markers, almost as if each streetlight bore witness to a theft. There’s a sharp concentration near transit hubs and popular neighborhoods, suggesting a correlation between foot traffic, bike use, and crime opportunity. The City has been making efforts to combat this issue; for example, Cycle Toronto launched the #EndBikeTheft campaign, advocating for secure parking and better registration systems, with city discussions exploring collaboration with platforms like 529 Garage to help reduce and recover stolen bikes (@citeEndBikeTheft). It’s almost as though you can hear the honks and feel the rush of people in these crowded streets.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Load necessary libraries
library(ggplot2)
library(sf)
library(osmdata)
library(here)

# Load cleaned coordinates data
coordinates_data <- read.csv(here("data/cleaned_coordinates_data.csv"))

# Get bounding box for Toronto (approximate limits for Toronto)
toronto_bbox <- c(-79.6303856025883, 43.5822069220627, -79.1180336515019, 43.8554571861712)

# Fetch map data from OpenStreetMap (OSM)
toronto_map <- opq(bbox = toronto_bbox) %>%
  add_osm_feature(key = "highway", value = "primary") %>%
  add_osm_feature(key = 'highway', value = 'secondary') %>%
  add_osm_feature(key = 'highway', value = 'tertiary') %>%
  add_osm_feature(key = 'highway', value = 'residential') %>%
  osmdata_sf()

# Convert map data to Simple Features (sf) objects
streets <- toronto_map$osm_lines

# Plot the map with bike theft locations data
bike_theft_plot <- ggplot() +
  geom_sf(data = streets, color = "gray", size = 0.3, alpha = 0.6) + # OSM street map layer
  geom_point(
    data = coordinates_data, aes(x = longitude, y = latitude),
    size = 0.5, color = "red"
  ) + # Bike theft points
  labs(
    title = "Map of Stolen Bike Cases in Toronto",
    x = "Longitude", y = "Latitude"
  ) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  theme(
    panel.background = element_blank(), # Remove panel background
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

bike_theft_plot
```

Finally, the third chart takes a step back to help us understand the data by neighborhood, but this time with broader strokes. It presents different regions in Toronto as bubbles of varying sizes and colors, indicating the number of thefts in each. The larger and darker the bubble, the more incidents that have occurred there. One particular neighborhood stands out—a massive blue circle in the downtown area—it visually reinforces what the other charts have hinted at: that specific neighborhoods, especially those near the core, are the epicenters of bicycle theft. Meanwhile, the areas with smaller circles seem almost untouched in comparison, a stark contrast that reveals a city of disparities—some areas buzz with activity and theft risk, while others enjoy relative calm.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#### Map with Areas ####

# Load necessary libraries
library(dplyr)
library(readr)
library(ggplot2)
library(plotly)
library(sf)
library(crosstalk)
library(here)
library(osmdata)

# Load your dataset (adjust the file path accordingly)
map_data <- read_csv(here("data", "map_by_area_data.csv"), show_col_types = FALSE)

# Get bounding box for Toronto (approximate limits for Toronto)
toronto_bbox <- c(-79.6303856025883, 43.5822069220627, -79.1180336515019, 43.8554571861712)

# Fetch map data from OpenStreetMap (OSM)
toronto_map <- opq(bbox = toronto_bbox) %>%
  add_osm_feature(key = "boundary", value = "administrative") %>%
  osmdata_sf()

# Convert map data to Simple Features (sf) objects
streets <- toronto_map$osm_lines

# Generate a similar plot using ggplot2 and plotly
shared_data <- crosstalk::SharedData$new(map_data)

# Create ggplot object with the dataset and Toronto map as a background
ggplot_map <- ggplot() +
  geom_sf(data = streets, color = "lightgray", size = 0.3, alpha = 0.6) +  # Add OSM street map layer
  geom_point(data = shared_data, aes(x = longitude, y = latitude, color = case_num, size = case_num, 
                                     text = paste("Total Cases: ", case_num, " | Area Name: ", area_name))) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  labs(title = "Toronto Stolen Bike Case Numbers by Area", x = "Longitude", y = "Latitude", color = "Number of Cases", size = "Total Cases") +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_color_continuous(labels = scales::comma) +
  scale_size_continuous(range = c(3, 12)) # Adjust range values to control the size of points

# Convert ggplot object to interactive plotly object
plotly_map <- ggplotly(ggplot_map, tooltip = "text") %>%
  highlight(
    on = "plotly_hover",
    off = "plotly_doubleclick",
    selected = attrs_selected(line = list(color = "black"))
  )

plotly_map
```

# Discussion 
- Consolidation of all the info that we present, we first talk about...this is how all visualizations are connected together, in the future what can be done...

# Conclusion
- short reminder for the audience, tips to protect their bikes

General tips: 
 - This paper wants to tell the story about tips to protect bikes

 - after every visualizations talk a bit about the visualization itself, key points, make sure the story that we are talking about is proved by this visualization so each plot make sense (for example in 2016 there is a peak in May illustrated by the visualization, why? find relevant news and articels to explain it and cite), 用更descriptive的形式去explain 你的visualization plot with proves.

# Citation