### 2024-11-19 paper/paper.Rmd
rewitten in story format

---
output: html_document
thanks: "Code and data are available at: [https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main](https://github.com/iJustinn/Toronto_Bicycle_Thefts/tree/main)."
---

<style>
/* Container for title, subtitle, description, author, and date with background image */
.header-section {
  background-image: url('bikes.jpg'); 
  background-size: cover;
  background-position: center;
  background-attachment: fixed;
  padding: 50px;
  color: white;
}

.title {
  font-size: 40px;                   /* Larger font size for the main title */
  font-weight: bold;                 /* Bold font weight for emphasis */
  color: #FFC107;                    /* Yellow color similar to the example */
  text-align: left;                  /* Align text to the left */
  font-family: Georgia, serif;       /* Use Georgia or a similar serif font */
  margin: 0;
  line-height: 1.1;                  /* Tighter line height */
  padding: 10px;                     /* Add padding around text */
  border: 5px solid #FFC107;         /* Border around the title text */
  border-radius: 8px;                /* Rounded corners for a softer look */
  display: inline-block;             /* Keeps the border tight around text */
  text-shadow: 3px 3px 8px rgba(0, 0, 0, 0.7); /* Text shadow for emphasis */
  box-shadow: 0px 4px 10px rgba(0, 0, 0, 0.5); /* Optional shadow for depth */
}

.subtitle {
  font-size: 28px;            /* Slightly larger font size for the subtitle */
  font-weight: normal;        /* Normal font weight */
  color: white;               /* White color for the subtitle */
  text-align: left;           /* Align text to the left */
  font-family: 'Times New Roman', Times, serif; /* A serif font for contrast */
  margin-top: 10px;
  line-height: 1.2;           /* Adjust line height for spacing */
  text-shadow: 3px 3px 8px rgba(0, 0, 0, 0.7); /* Text shadow for emphasis */
}

.description {
  display: block; /* Makes span act like a block element */
  font-size: 18px;
  font-weight: normal;
  font-style: italic; /* Makes the text italic */
  line-height: 1.6;
  text-align: left;
  max-width: 600px;
  margin-top: 20px;
  font-family: "Times New Roman", Times, serif; /* Change to desired font */
  background-color: transparent !important;
  border: none !important;
  padding: 0 !important;
  box-shadow: none !important;
  text-shadow: 3px 3px 8px rgba(0, 0, 0, 0.7); /* Text shadow for emphasis */
}

.author-date {
  font-size: 16px;
  font-weight: normal;
  color: white;
  text-align: left;
  margin-top: 30px;
  font-style: italic;
  text-shadow: 3px 3px 8px rgba(0, 0, 0, 0.7); /* Text shadow for emphasis */
}

.main-body {
  max-width: 800px;              /* Set max width for main body */
  width: 100%;                    /* Ensures full width within max-width */
  font-size: 12px;                /* Font size for readability */
  font-weight: 300;               /* Lighter font weight for softer tone */
  color: black;             
  line-height: 1.5;               /* Increase line height for readability */
  text-align: justify;            /* Justified alignment for a clean look */
  margin: 0 auto;                 /* Center content on the page */
  padding: 20px;                  /* Add padding around the text */
  font-family: 'Georgia', serif;  /* Use a classic serif font */
}

.main-body p {
  width: 100%;                    /* Full width within .main-body */
  max-width: none;                /* Remove max-width constraint */
  margin: 0 0 1em 0;              /* Consistent margin below each paragraph */
  padding: 0;                     /* Remove padding */
}

.section-heading {
  max-width: 800px;              /* Limit the width */
  margin: 20px auto 20px auto;   /* Center the heading on the page */
  padding-bottom: 5px;           /* Space below for the underline */
  border-bottom: 2px solid #FFC107; /* Optional yellow underline */
}

.section-title-line1 {
  font-size: 25px;               /* Larger font size for the first line */
  font-weight: bold;             /* Bold weight for emphasis */
  color: #FFC107;                /* Dark yellow color */
  text-align: left;              /* Align text to the left */
  font-family: 'Georgia', serif; /* Serif font */
  display: block;                /* Ensures each line appears separately */
}

.section-title-line2 {
  font-size: 20px;               /* Slightly smaller font for the second line */
  font-weight: normal;           /* Regular weight */
  color: #603601;                /* Dark color for contrast */
  text-align: left;              /* Align text to the left */
  font-family: 'Georgia', serif; /* Serif font */
  display: block;                /* Ensures each line appears separately */
  margin-top: -5px;               /* Adds some spacing between lines */
}

/* Citation section styling */
.citation-section {
  max-width: 800px;
  margin: 30px auto;              /* Center the section */
  padding: 20px;                  /* Add padding around citations */
  background-color: #f9f9f9;      /* Light background for contrast */
  border-left: 5px solid #FFC107; /* Yellow border to match title color */
  box-shadow: 0px 4px 10px rgba(0, 0, 0, 0.1); /* Subtle shadow for depth */
  font-family: 'Georgia', serif;  /* Match main font */
  color: #333;                    /* Dark text color */
}

.citation-section h3 {
  font-size: 20px;                /* Set citation title font size to 20px */
  line-height: 1.5;
  color: #603601;                 /* Darker color for contrast */
  font-weight: bold;
  margin-bottom: 10px;
}

.citation-section ul {
  list-style-type: none;          /* Remove default bullets */
  padding: 0;                     /* Remove default padding */
}

.citation-section li {
  font-size: 12px;                /* Set individual citations to 12px */
  margin-bottom: 10px;            /* Space between each citation */
  line-height: 1.5;               /* Increase line height for readability */
}

.citation-section li a {
  color: #1a73e8;                 /* Link color for readability */
  text-decoration: none;
}

.citation-section li a:hover {
  text-decoration: underline;      /* Underline on hover */
}

</style>

<!-- Header section with background image -->
<div class="header-section">
  <div class="title">Ride at Your Own Risk: Insights into Toronto’s Bicycle Theft Hotspots</div>
  <div class="subtitle">Navigating Toronto’s streets with eyes wide open, this data-driven guide helps cyclists stay ahead of bicycle thieves.</div>

  <!-- Using span instead of div for description -->
  <span class="description">
    Each year, countless bikes disappear from Toronto’s streets, leaving cyclists frustrated and vulnerable. Our analysis digs into patterns and insights across the city, highlighting where and when bicycles are most at risk. With data spanning over a decade, we reveal theft hotspots and provide tips to help keep your bike safe.
  </span>

  <div class="author-date">Yingke He, Ziheng Zhong<br>November 18, 2024</div>
</div>

<!-- Main article content without background image -->
<div class="main-body">
  <p>Imagine locking up your bike on a bustling Toronto street, feeling reassured that a strong lock and some common sense will keep it safe. But what if certain seasons, specific bike types, or even certain neighborhoods increase your risk of theft far more than you realize? For many cyclists, bike theft has moved from a distant possibility to an everyday concern, with data revealing just how pervasive the risk can be in Toronto.</p>

  <p>In this analysis, we dig into over a decade’s worth of data from the <a href="https://data.torontopolice.on.ca/datasets/TorontoPS::bicycle-thefts-open-data/about" target="_blank" style="color: #FFC107; text-decoration: underline;">Toronto Police Service</a> (Toronto Police Service, 2024c) to answer key questions: When are bikes at the greatest risk? What makes a bike more likely to be targeted? And where are the theft hotspots cyclists should avoid? Each section of our article addresses a different angle of bike theft in Toronto, from seasonal and weekly theft patterns to variations by bicycle type and value, to pinpointing the city’s most high-risk locations.</p>

  <p>Over the past decade, <strong style="color: #FFC107;">32,489</strong> bike thefts have been reported in Toronto, yet only <strong style="color: #FFC107;">422</strong> of these stolen bikes were successfully recovered. With these insights, we aim to empower cyclists with knowledge that can help them safeguard their bikes and inform Cycle Toronto and local authorities about the areas most in need of intervention. Read on as we trace theft trends over the years, reveal the high-risk locations, and provide practical advice to help protect your bike.</p>
</div>

<div class="main-body">
  <p>Hi there! I’m a bicycle—just your regular two-wheeled friend. I’m here to take you places, let you feel the wind in your hair, and help you enjoy the streets of Toronto. Life might seem easy for me, but it’s not all smooth rides. In this city, there’s one big worry I face all the time: bike theft. And believe me, I’ve seen plenty of it.</p>

  <p>Picture this: you lock me up on a busy Toronto street, thinking a strong lock and a little care will keep me safe. But the truth is, some seasons, certain types of bikes, and even specific areas make bikes like me much more likely to be stolen. Over the years, bike theft has gone from something that “might happen” to something we constantly have to worry about. But don’t worry—I’ve joined forces with some data experts to figure out what’s really happening.</p>

  <p>We’ve looked at over ten years of data from the <a href="https://data.torontopolice.on.ca/datasets/TorontoPS::bicycle-thefts-open-data/about" target="_blank" style="color: #FFC107; text-decoration: underline;">Toronto Police Service</a> (Toronto Police Service, 2024c) to answer some big questions: When are bikes like me at the most risk? What makes us targets? And which parts of Toronto should we avoid? Over this time, <strong style="color: #FFC107;">32,489</strong> bikes have been reported stolen in Toronto, but only <strong style="color: #FFC107;">422</strong> were ever recovered. With these insights, we hope to give cyclists like you the knowledge to protect us better, while helping groups like Cycle Toronto and local officials focus on areas that need the most attention. Together, we’ll uncover theft trends, identify high-risk locations, and share practical tips to keep your bike safe and secure.</p>
</div>

```{r, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, include=TRUE, eval=TRUE)

required_packages <- c("sf", "here", "dplyr", "readr", "purrr", "scales", "plotly", "ggplot2", "leaflet", "osmdata", 
                       "ggridges", "cowplot", "lubridate", "tidyverse", "crosstalk", "bookdown", "hrbrthemes", "gganimate")

for (p in required_packages) {
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(sf)
library(here)
library(dplyr)
library(readr)
library(purrr)
library(scales)
library(plotly)
library(ggplot2)
library(leaflet)
library(osmdata)
library(cowplot)
library(ggridges)
library(bookdown)
library(lubridate)
library(tidyverse)
library(crosstalk)
library(gganimate)
library(hrbrthemes)
library(htmltools)
```

<div class="section-heading">
  <span class="section-title-line1">When Are Bikes at the Greatest Risk?</span><br>
  <span class="section-title-line2">A Look at Seasonal and Weekly Patterns</span>
</div>

<center>
```{r interactive_weekly_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=8, fig.height=3.5}

# Load the data
data <- read_csv(here("data", "Theft_Counts_By_Day.csv"), show_col_types = FALSE)

# Arrange data by day of the week (Monday to Sunday) and convert OCC_DOW to a factor with specified levels
data <- data %>%
  mutate(OCC_DOW = factor(OCC_DOW, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  arrange(OCC_DOW)

# Initialize plot with a line trace for the entire data
fig <- plot_ly(data, x = ~OCC_DOW, y = ~Number_of_Thefts, type = 'scatter', mode = 'lines+markers',
               line = list(color = '#05595B'), marker = list(color = '#603601', size = 10))

# Add points for each day with initial visibility settings
for (i in 1:nrow(data)) {
  fig <- fig %>%
    add_trace(
      x = data$OCC_DOW[i],
      y = data$Number_of_Thefts[i],
      type = 'scatter',
      mode = 'markers+text',
      text = paste("Thefts:", data$Number_of_Thefts[i]),
      textposition = "top center",
      marker = list(color = "#FFC107", size = 15),
      showlegend = FALSE,
      visible = ifelse(i == 1, TRUE, FALSE)
    )
}

# Define steps for each day in the slider
steps <- lapply(1:nrow(data), function(i) {
  step <- list(
    method = "restyle",
    args = list("visible", as.list(rep(FALSE, nrow(data) + 1))),  # +1 to include the main line trace
    label = data$OCC_DOW[i]
  )
  step$args[[2]][i + 1] <- TRUE  # Make only the current day’s point visible (offset by 1 due to main line trace)
  step$args[[2]][1] <- TRUE       # Keep the main line trace visible
  return(step)
})

# Add layout with the slider, title annotation in the top-left, and set background colors
fig <- fig %>%
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Number of Thefts"),
    sliders = list(
      list(
        active = 0,
        currentvalue = list(prefix = "Day: "),
        steps = steps
      )
    ),
    plot_bgcolor = "#faecd5",     # Set background color for the plotting area
    paper_bgcolor = "#faecd5",     # Set background color for the entire plot
    annotations = list(
      list(
        x = 0.01, y = 1.05,  # Position in normalized coordinates (0 = left, 1 = top)
        xref = "paper", yref = "paper",
        text = "Weekly Bicycle Theft Patterns in Toronto",  # Title text
        showarrow = FALSE,
        font = list(size = 11, color = "#603601")  # Set title color to #603601
      )
    )
  )

# Display the plot
fig

``` 
</center>
<div style="color: grey; font-size: 12px; margin-top: 3px; max-width: 800px; margin: 0 auto;">
  Note: Dragging the slider reveals daily bike theft patterns in Toronto, highlighting higher-risk days, with data covering 2014 to mid-2024.
</div> \

<div class="main-body">
<p>It was Friday again, my least favorite day of the week. Don’t get me wrong, I love being ridden to the park or to the office, feeling the wind rushing past my handlebars. But Fridays? Fridays are dangerous. I’ve seen too many of my fellow bikes disappear on this day—snatched from racks, left lonely in crowded parking lots. Over <strong style="color: #FFC107;">4,900</strong> of us have gone missing on Fridays in Toronto. It’s not paranoia; it’s data.</p>

<p>You see, Fridays are chaotic. People rush to wrap up their week, juggling work and plans for the weekend. Some leave us in high-traffic zones, near pubs or downtown hotspots, forgetting how vulnerable we are. By Saturday and Sunday, the risk eases up a bit. Fewer commutes, fewer hurried stops. But Fridays? Fridays are a minefield.</p>

<p>If only they knew what the data shows. Advocacy groups like <a href="https://www.cycleto.ca/" target="_blank" style="color: #FFC107; text-decoration: underline;">Cycle Toronto</a> (Cycle Toronto, 2021) have been raising awareness about these patterns. Perhaps one day, local authorities will patrol more on Fridays or run public awareness campaigns around hotspots. And for cyclists, knowing that Friday is the peak day for thefts might encourage them to lock us more securely or park in safer areas. I just hope these efforts take hold before more of us end up as another Friday statistic.</p\ 
</div>

<center>
```{r mothly_plot, include = TRUE,fig.cap = "Bike theft and recovery patterns in Toronto from 2014 to 2024", warning = FALSE, message = FALSE, fig.width=8, fig.height=4.3}

# Load your data
data <- read_csv(here("data", "Theft_and_Found_Counts_By_Month_2014_2024.csv"), show_col_types = FALSE)

# Convert month names to month numbers for ordering
data <- data %>%
  mutate(OCC_MONTH_NUM = factor(OCC_MONTH, levels = month.name))

# Specify levels for OCC_YEAR to order from 2014 to 2024
data$OCC_YEAR <- factor(data$OCC_YEAR, levels = 2014:2024)

# Calculate a scaled height for "Found" data by dividing by the maximum "Found" value
max_found <- max(data$Found, na.rm = TRUE)
max_stolen <- max(data$Stolen, na.rm = TRUE)

# Calculate yearly totals for annotation
yearly_totals <- data %>%
  group_by(OCC_YEAR) %>%
  summarise(Found_Total = sum(Found, na.rm = TRUE), Stolen_Total = sum(Stolen, na.rm = TRUE))

# Plot for Bikes Found with scaled heights and yearly total annotations
plot_found <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                               height = (Found / max_found) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#05595B", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    plot.title = element_text(size = 13, color = "#05595B", hjust = -0.1, vjust = 1.5),  # Smaller title
    panel.background = element_rect(fill = "#faecd5", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#faecd5", color = NA)    # Light blue background for overall plot
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Found") +  # Add title for Bikes Found
  geom_text(data = yearly_totals, aes(x = 12, y = OCC_YEAR, label = Found_Total), 
            color = "#05595B", hjust = 1, size = 3, inherit.aes = FALSE)

# Plot for Bikes Stolen with scaled heights, no borders, and yearly total annotations
plot_stolen <- ggplot(data, aes(x = OCC_MONTH_NUM, y = OCC_YEAR, 
                                height = (Stolen / max_stolen) * 4, group = OCC_YEAR)) +
  geom_ridgeline(
    fill = "#603601", color = "white", alpha = 0.5, scale = 1, linewidth = 0
  ) +
  labs(y = "", x = NULL) +  # Remove x-axis title
  theme_minimal() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(size = 0.5),
    panel.border = element_blank(),
    panel.background = element_rect(fill = "#faecd5", color = NA),  # Light blue background for plot area
    plot.background = element_rect(fill = "#faecd5", color = NA),   # Light blue background for overall plot
    plot.title = element_text(size = 13, color = "#603601", hjust = -0.1, vjust = 1.5)  # Smaller title
  ) +
  scale_x_discrete(labels = month.abb) +
  ggtitle("Bikes Stolen") +  # Add title for Bikes Stolen
  geom_text(data = yearly_totals, aes(x = 12, y = OCC_YEAR, label = Stolen_Total), 
            color = "#603601", hjust = 1, size = 3, inherit.aes = FALSE)

# Combine the plots side by side
combined_plot <- plot_grid(plot_found, plot_stolen, nrow = 1)

# Display the combined plot
combined_plot

```
</center>
\

<div class="main-body">
  <p>Summer is supposed to be my favorite time of year. The warm weather, the long days, and the bustling streets filled with fellow bikes and riders. But as much as I love the attention, it also makes me nervous. Between May and August, thefts peak—this is when bikes like me are at the greatest risk. It’s not surprising, really. More people are out riding, whether it’s for fun or as their main way to get around, as the <a href="https://www.toronto.ca/wp-content/uploads/2021/04/8f76-2019-Cycling-Public-Option-Survey-City-of-Toronto-Cycling.pdf" target="_blank" style="color: #FFC107; text-decoration: underline;">City of Toronto</a> (City of Toronto, 2019) has pointed out. Events like <a href="https://www.bikemonth.ca/toronto" target="_blank" style="color: #FFC107; text-decoration: underline;">Bike Month</a> (Bike Month, 2024) in June bring even more of us out into the open, where thieves have more opportunities to strike. We become easy targets, exposed and vulnerable, even when we’re locked up securely.</p>

  <p>And then there’s the sad reality of being found. While thousands of us go missing every year, the number of bikes recovered stays painfully low, no matter the season. Despite efforts like the <a href="https://www.tps.ca/services/bicycle-registration/" target="_blank" style="color: #FFC107; text-decoration: underline;">city-wide bike registry programs</a> (Toronto Police Service, 2024a) and <a href="https://www.tps.ca/crime-prevention/bicycle-safety-and-theft-prevention/" target="_blank" style="color: #FFC107; text-decoration: underline;">police campaigns</a> (Toronto Police Service, 2024b) to teach better locking techniques, the chances of being reunited with my rider are slim. Even with new <a href="https://www.toronto.ca/services-payments/streets-parking-transportation/cycling-in-toronto/bicycle-parking/" style="color: #FFC107; text-decoration: underline;">bike parking programs</a> (City of Toronto, 2024) and secure lock stands, the efforts still feel small compared to how many of us disappear each year. I can’t help but wonder—what more can be done to keep me and my friends safe?</p>
</div>

<div class="section-heading">
  <span class="section-title-line1">What Makes a Bike a Target?</span><br>
  <span class="section-title-line2">Analyzing Theft by Bicycle Type and Value</span>
</div>

<center>
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=7, fig.height=4}

# Load the cleaned dataset
bike_type_counts <- read_csv(here("data", "bike_type_counts.csv"))

# Mapping bike type codes to descriptive names
bike_type_names <- c(
  "Others" = "Other Types",
  "MT" = "Mountain",
  "RG" = "Gravel",
  "OT" = "Other",
  "RC" = "Recreational & Commuting",
  "EL" = "Electric"
)

# Calculate percentages for each bike type
bike_type_counts <- bike_type_counts %>%
  mutate(total = sum(bike_num)) %>%
  group_by(bike_type) %>%
  mutate(percentage = (bike_num / total) * 100)

# Summarize types with less than 5% as "Others"
bike_type_counts <- bike_type_counts %>%
  mutate(bike_type = ifelse(percentage < 5, "Others", bike_type)) %>%
  group_by(bike_type) %>%
  summarise(bike_num = sum(bike_num), percentage = sum(percentage)) %>%
  ungroup()

# Reorder bike types by percentage in descending order
bike_type_counts <- bike_type_counts %>%
  mutate(bike_type = fct_reorder(bike_type, -percentage))

# Create a horizontal stacked bar chart
ggplot(bike_type_counts, aes(x = factor(1), y = percentage, fill = bike_type)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_fill(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Distribution of Bike Types",
    x = "Bike Types",
    y = "Percentage",
    fill = "Bike Types"
  ) +
  scale_fill_manual(labels = bike_type_names, values = c("#FF7043", "#FF8F00", "#FFA726", "#FFB300", "#FFC107", "#FFCA28")) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "right")

```
</center>

<div class="main-body">
  <p>Let me tell you about the bikes that thieves seem to like the most. It’s no surprise that mountain bikes like me and our commuting cousins top the list. We’re everywhere—on trails, on streets, parked outside offices, and near parks. Together, we make up the biggest targets, with mountain bikes accounting for <strong style="color: #FFC107;">31%</strong> and recreational & commuting bikes adding another <strong style="color: #FFC107;">27.4%</strong>. Our popularity means we’re easy to spot and easier to steal. </p> <p>Electric bikes, though, seem to have it a bit easier. They only make up <strong style="color: #FFC107;">7.8%</strong> of stolen bikes. Maybe it’s because their owners tend to invest more in security—fancier locks, GPS trackers, or even parking us in safer spots. I guess it pays to be high-tech, but for the rest of us, being useful and everywhere means we’re always at risk.</p>

<center>
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=7, fig.height=4}

# Load data
bike_cost_counts <- read_csv(here("data", "bike_cost_counts.csv"))

# Create a pie chart of bike cost tiers
ggplot(bike_cost_counts, aes(x = "", y = bike_num, fill = cost_tier)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y") +
  labs(
    title = "Proportion of Bike Cost Tiers",
    fill = "Cost Tier"
  ) +
  scale_fill_manual(values = c("#FFCA28", "#FFC107", "#FFB300", "#FF8F00")) +  # Customize colors
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5)
  ) +
  geom_text(aes(label = paste0(round(bike_num / sum(bike_num) * 100, 1), "%")),
            position = position_stack(vjust = 0.5))

```
</center>

  <p>Let’s talk about how much I’m worth—or, as thieves might see it, how much they think I’m worth. The second plot shows that bikes in the lower price ranges, like those under \$500 or between \$500-\$999, are stolen the most, making up <strong style="color: #FFC107;">34.9%</strong> and <strong style="color: #FFC107;">31.7%</strong> of thefts. It makes sense—we’re often used for daily commuting, left outside offices, stores, or train stations, and become easy pickings. But don’t think being fancy makes me immune. Bikes worth \$1000-\$1999 and even those over \$2000 aren’t safe either, accounting for <strong style="color: #FFC107;">20.9%</strong> and <strong style="color: #FFC107;">12.4%</strong> of thefts. In the city, no bike is truly off-limits.</p>

  <p>What does this mean for me and my friends? It’s simple: we all need protection, no matter our price tag. For lower-cost bikes like me, a strong U-lock or chain lock is a must. Parking in busy spots with security cameras or near bike racks helps too—anything to make thieves think twice. Even if I’m not worth a lot of money, taking these steps can make a big difference in keeping me safe.</p>

  <p>Higher-value bikes, though, need even more care. If I were a high-end model, I’d hope for two locks—maybe a U-lock and a cable lock—and I’d beg you not to leave me outside overnight. Indoor storage is the gold standard. Oh, and registering me with Toronto’s bike registry or <a href="https://project529.com" target="_blank" style="color: #FFC107; text-decoration: underline;">Project 529</a> (Toronto Police Service, 2024b; Project 529, 2024) would improve my chances of being found if something went wrong. Marking me with a unique ID or changing up where you park me can also keep thieves guessing. With the right precautions, we can stay out of trouble and keep riding the streets of Toronto together.</p>
</div>

<div class="section-heading">
  <span class="section-title-line1">A Decade of Disappearing Bikes:</span><br>
  <span class="section-title-line2"> Tracing Theft Trends Over Ten Years</span>
</div>

<center>
```{r animated_bar_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=4, fig.height=2}

# Load the data
data <- read_csv(here("data", "Stolen_Found_Counts.csv"), show_col_types = FALSE)

# Rename categories for clarity
data <- data %>%
  mutate(Offence_Category = case_when(
    Offence_Category == "THEFT" ~ "Bikes Stolen",
    Offence_Category == "FOUND" ~ "Bikes Found",
    TRUE ~ Offence_Category
  ))

# Set the number of frames for the animation
max_frames <- 100  # Total number of frames for smooth animation

# Generate interpolated data for simultaneous increase
animated_data <- data %>%
  # Repeat each category for each frame
  group_by(Offence_Category) %>%
  do(data.frame(
    Offence_Category = .$Offence_Category,
    Number_of_Offences = seq(0, .$Number_of_Offences, length.out = max_frames),
    frame = 1:max_frames
  ))

# Ensure Offence_Category is a factor for proper ordering
animated_data$Offence_Category <- factor(animated_data$Offence_Category, levels = c("Bikes Found", "Other", "Bikes Stolen"))

# Define custom colors for each category
custom_colors <- c("Bikes Stolen" = "#603601", "Bikes Found" = "#4682B4", "Other" = "#eedaaf")

# Create the animated bar plot with the x-axis label at the bottom right
ggplot(animated_data, aes(x = Offence_Category, y = Number_of_Offences, fill = Offence_Category)) + 
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Offense Category', y = 'Number of Offences') +

  # Apply custom colors
  scale_fill_manual(values = custom_colors) +

  # Remove the legend title
  guides(fill = guide_legend(title = NULL)) +

  # Set background colors and adjust text positioning, placing the x-axis label at the bottom right
  theme(
    panel.background = element_rect(fill = "#faecd5", color = NA),
    plot.background = element_rect(fill = "#faecd5", color = NA),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.x = element_text(size = 6, hjust = 1, vjust = -1),  # Position x-axis label at the bottom right
    axis.title.y = element_text(size = 6),
    legend.text = element_text(size = 4)
  ) +

  # gganimate specific settings for smooth transition
  transition_states(
    frame,
    transition_length = 2,
    state_length = 1
  ) +
  ease_aes('linear')  # Ensures a smooth, even increase

```
</center>

<div style="color: grey; font-size: 12px; margin-top: 3px; max-width: 800px; margin: 0 auto;">
  Displays the count of various bicycle-related offences reported to the Toronto Police Service, highlighting categories such as stolen, found, and other incidents. The smooth increase in bar heights emphasizes the scale of each offence type, offering insights into the distribution and relative frequency of these incidents from 2014 to 2024.
</div> \

<div class="main-body">
  <p>Now move on to the next plot, it’s hard to miss the tallest bar on the chart. It represents bikes like me—stolen over <strong style="color: #FFC107;">32,000</strong> times across Toronto. That’s a staggering number and a clear sign of just how common bike theft has become. Hotspots like downtown Toronto and areas near major transit hubs are especially risky, as pointed out by <a href="https://www.ctvnews.ca/canada/should-toronto-tear-up-its-bike-lanes-to-improve-traffic-flow-critics-say-it-s-not-so-simple-1.7103768" style="color: #FFC107; text-decoration: underline;">CTV News</a> (CTV News, 2024). Cyclists often leave us unattended there, making it easier for thieves to strike. Thankfully, programs like <a href="https://www.tps.ca/media-centre/news-releases/fix-title-project-sentry/" style="color: #FFC107; text-decoration: underline;">Project Sentry</a> (Toronto Police Service, 2024d) are trying to change that. Officers in 14 Division are working hard to recover bikes like me and reunite us with our owners. It’s a step in the right direction, but there’s still a long way to go.</p>

  <p>Then there’s the “Bikes Found” bar, standing sadly at just <strong style="color: #FFC107;">422</strong>. It’s tiny compared to the number of stolen bikes, showing how difficult it is for me and my friends to find our way back home. Even with tools like <a href="https://project529.com/garage" style="color: #FFC107; text-decoration: underline;">Project 529</a> (Project 529, 2024), which helps track bike ownership, recovery is a real challenge. Sometimes, thieves change how we look or sell us off, making it hard for authorities to identify us. With limited resources, finding us often feels like searching for a needle in a haystack.</p>

  <p>And let’s not forget the “Other” category, with about <strong style="color: #FFC107;">3,200</strong> cases. This includes all sorts of problems—attempted thefts, damage, mischief, and more. Even if we’re not stolen outright, it’s a reminder that staying safe in Toronto is tough. Whether it’s scratches, broken parts, or near-misses, we face plenty of challenges out here on the streets.</p>
</div>

<center>
```{r animated_line_plot, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=4.2, fig.height=2.6}

# Load your data
data <- read_csv(here("data", "Bicycle_Theft_Monthly_Counts.csv"), show_col_types = FALSE)

# Create a numeric month order variable
data <- data %>%
  mutate(
    OCC_MONTH_NUM = match(OCC_MONTH, c("January", "February", "March", "April", "May", 
                                       "June", "July", "August", "September", "October", 
                                       "November", "December"))
  )

# Create the animated plot with smaller text elements
plot <- data %>%
  ggplot(aes(x = OCC_MONTH_NUM, y = Number_of_Thefts, group = as.factor(OCC_YEAR), color = as.factor(OCC_YEAR))) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d() +
  ylab("Number of Bikes Stolen") +
  xlab("Month") +
  scale_x_continuous(breaks = 1:12, labels = c("January", "February", "March", "April", "May", "June", 
                                               "July", "August", "September", "October", "November", "December")) +
  theme_ipsum() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 8),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 6),
    panel.background = element_rect(fill = "#faecd5", color = NA),
    plot.background = element_rect(fill = "#faecd5", color = NA)
  ) +
  transition_reveal(OCC_MONTH_NUM) +
  labs(color = "Year")

# Render the animation
animate(plot, nframes = 100, fps = 10)

```
</center>

<div style="color: grey; font-size: 12px; margin-top: 3px; max-width: 800px; margin: 0 auto;">
 Visualizes the monthly count of bicycle thefts reported to the Toronto Police Service from 2014 to June 2024. The animation reveals seasonal patterns and variations across different years, providing insights into peak months and potential trends over time.
</div>

<div class="main-body">
  <p>Every year, like clockwork, we bikes notice the change in seasons. The animated line plot shows it perfectly—how thefts rise with the spring sunshine, peak in the summer months of June and July, and then ease up as the colder weather sets in. It’s no surprise; during the warmer months, we’re everywhere. Torontonians take us out for daily commutes, weekend rides, and everything in between. But this also means we’re more exposed, left in public spaces where thieves can easily find us. Summer, for us, is both a season of joy and of danger, a time when heightened security measures are absolutely necessary.</p>

  <p>Then came 2020—a year none of us will ever forget. The usual seasonal rhythm was disrupted by the COVID-19 pandemic. With public transit suddenly seen as risky, more people turned to us for safer, flexible transportation. Bike ownership soared, and with it, unfortunately, thefts did too (Shapiro, 2021). At the same time, police resources were stretched thin, and patrols or recovery efforts likely took a backseat. It was a tough time to be a bike in Toronto. This spike in thefts during 2020 reminds us that external factors, like a public health crisis, can make an already challenging situation worse. It shows how important it is to adapt theft prevention strategies to changing urban needs so we can keep rolling safely, no matter what the future holds.</p>
</div>

<div class="section-heading">
  <span class="section-title-line1">Mapping the Danger Zones:</span><br>
  <span class="section-title-line2"> High-Risk Locations Across Toronto</span>
</div>

<div class="main-body">
  <p>Now, imagine being me, trying to figure out where I’m safest on the streets of Toronto. The interactive map below feels like a glimpse into my world—bright orange circles marking the spots where bikes like me are most at risk. Each circle tells a story: the smaller ones mean fewer incidents, while the large, glaring ones scream danger. And then there’s <a href="https://en.wikipedia.org/wiki/Downtown_Toronto" target="_blank" style="color: #FFC107; text-decoration: underline;">Downtown Toronto</a>, the heart of the city, covered by a massive bubble. It’s no surprise—this bustling area, filled with commuters and visitors, is the epicenter of bike thefts. As you look further north and into suburban areas, the bubbles shrink, suggesting quieter streets where bikes are often stored more securely.</p>

  <p>What makes this map even better? You can <strong style="color: #FFC107;">click</strong> on the bubbles to zoom in. It’s like having a detective’s guide to uncover the details—exact locations, hot spots, and even the safer neighborhoods. For me and my fellow bikes, this kind of insight is priceless. Knowing where the risks are helps us stay out of trouble and keep rolling safely through Toronto’s streets.</p>
</div>

<center>
```{r interactive_map, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=7, fig.height=4.5}
#### Map with Markers ####

# Load your dataset
map_data <- read_csv(here("data", "map_by_neighbourhoods_data.csv"), show_col_types = FALSE)

# Modify the data frame, categorizing case_num column
map_data <- map_data %>% 
  mutate(
    mag.level = cut(case_num, c(100, 500, 1000, 1500),
                    labels = c('> 100 & <=500', '>500 & <=1000', '>1000 & <=1500'))
  ) %>%
  split(.$mag.level)

# Initialize the leaflet map
l <- leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap)

# Add the markers to the map for each case_num level
names(map_data) %>%
  walk(function(df) {
    l <<- l %>% 
      addMarkers(
        data = map_data[[df]], lng = ~longitude, lat = ~latitude,
        label = ~paste0("Hood: ", hood_name, " | Stolen Place: ", stolen_place, " | Cost: ", cost, " | Status: ", status),
        popup = ~paste0(
          "Hood: ", hood_name, "<br>",
          "Stolen Place: ", stolen_place, "<br>",
          "Cost: ", cost, "<br>",
          "Status: ", status, "<br>",
          "Case in area: ", case_num
        ),
        group = df,
        clusterOptions = markerClusterOptions()
      )
  })

# Add layer controls and a minimap
l <- l %>%
  addLayersControl(
    overlayGroups = names(map_data),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addMiniMap(tiles = providers$Esri.OceanBasemap, width = 120, height = 80)

# Display the Map
l

```
</center>

<div style="color: grey; font-size: 12px; margin-top: 3px; max-width: 800px; margin: 0 auto;">
  Note: This interactive map reveals clusters of bicycle thefts across Toronto, clicking on each circle allows users to zoom in for detailed information on specific theft locations.
</div> \

<div class="main-body">
  <p>Take a broader look at the view of the city—it’s like flying on the sky for a little bit. A sea of <strong style="color: #FFC107;">yellow</strong> dots stretches across the map, each one marking where a bike like me was stolen (see "Map of Stolen Bike Cases in Toronto" below). In downtown Toronto, these dots are packed so tightly it looks like a thick yellow blanket covering the streets. It’s a clear reminder that the busy central parts of the city are the riskiest for us. Transit hubs and popular neighborhoods show the highest concentration, likely because they have the most people, bikes, and unfortunately, opportunities for theft.</p>

  <p>But not all hope is lost. Advocacy groups like Cycle Toronto have stepped up with initiatives like the <a href="https://www.cycleto.ca/council-talking-about-ending-bike-theft-week" target="_blank" style="color: #FFC107; text-decoration: underline;">#EndBikeTheft campaign</a> (Cycle Toronto, 2021), pushing for safer parking and better registration systems. There’s even talk of teaming up with platforms like <a href="https://project529.com/garage" style="color: #FFC107; text-decoration: underline;">529 Garage</a> (Project 529, 2024) to track us better and help us get back home if we’re ever taken. It’s a start, and with more efforts like these, maybe one day those yellow dots won’t cover the map quite so densely.</p>
</div>

<center>
```{r map_of_stolen_bike, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=8, fig.height=4}

# Load cleaned coordinates data
coordinates_data <- read.csv(here("data/cleaned_coordinates_data.csv"))

# Get bounding box for Toronto (approximate limits for Toronto)
toronto_bbox <- c(-79.6303856025883, 43.5822069220627, -79.1180336515019, 43.8554571861712)

# Fetch map data from OpenStreetMap (OSM)
toronto_map <- opq(bbox = toronto_bbox) %>%
  add_osm_feature(key = "highway", value = "primary") %>%
  add_osm_feature(key = 'highway', value = 'secondary') %>%
  add_osm_feature(key = 'highway', value = 'tertiary') %>%
  add_osm_feature(key = 'highway', value = 'residential') %>%
  osmdata_sf()

# Convert map data to Simple Features (sf) objects
streets <- toronto_map$osm_lines

# Plot the map with bike theft locations data
bike_theft_plot <- ggplot() +
  geom_sf(data = streets, color = "gray", size = 0.3, alpha = 0.6) + # OSM street map layer
  geom_point(
    data = coordinates_data, aes(x = longitude, y = latitude),
    size = 0.3, color = "#FFC107"
  ) + # Bike theft points
  labs(
    title = "Map of Stolen Bike Cases in Toronto",
    x = "Longitude", y = "Latitude"
  ) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  theme(
    panel.background = element_blank(), # Remove panel background
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

bike_theft_plot

```
</center>

<div style="color: grey; font-size: 12px; margin-top: 3px; max-width: 800px; margin: 0 auto; text-align: center; display: flex; justify-content: center;">
  This map reveals a high concentration of bike theft incidents in downtown Toronto.
</div>

<div class="main-body">
  <p>The map below gave us another step back, which is a chance for us to see how each areas compare when it comes to bike theft. It’s a colorful scene, with bubbles of different sizes and shades showing how many of us have been stolen in each area. The larger the bubble, the more incidents it represents. And right in the heart of the city is a massive <strong style="color: #FFC107;">yellow</strong> circle, covering downtown Toronto. It’s like a giant warning sign, confirming what we already know: the core of the city is the most dangerous place for bikes like me.</p>

  <p>Further out, the bubbles shrink, and the map starts to feel calmer. These smaller circles mark neighborhoods where thefts are less common, offering a glimpse of what safety might look like. It’s a sharp divide—some areas are hotspots, buzzing with theft activity, while others stay quiet and safer. For bikes like me, maps like this are a lifeline, helping us and our riders navigate Toronto’s streets with a better sense of where to stay extra cautious.</p>
</div>

<center>
```{r map_case_number, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=7, fig.height=4}
#### Map with Areas ####

# Load your dataset (adjust the file path accordingly)
map_data <- read_csv(here("data", "map_by_area_data.csv"), show_col_types = FALSE)

# Get bounding box for Toronto (approximate limits for Toronto)
toronto_bbox <- c(-79.6303856025883, 43.5822069220627, -79.1180336515019, 43.8554571861712)

# Fetch map data from OpenStreetMap (OSM)
toronto_map <- opq(bbox = toronto_bbox) %>%
  add_osm_feature(key = "boundary", value = "administrative") %>%
  osmdata_sf()

# Convert map data to Simple Features (sf) objects
streets <- toronto_map$osm_lines

# Generate a similar plot using ggplot2 and plotly
shared_data <- crosstalk::SharedData$new(map_data)

# Define more detailed breaks for color scale based on `case_num`
max_case_num <- max(map_data$case_num, na.rm = TRUE)
color_breaks <- seq(0, max_case_num, by = max_case_num / 4)

# Create a rounding function to simplify the labels
rounded_labels <- function(x) scales::comma(round(x, -6))

# Create ggplot object with the dataset and Toronto map as a background
ggplot_map <- ggplot() +
  geom_sf(data = streets, color = "lightgray", size = 0.3, alpha = 0.6) +  # Add OSM street map layer
  geom_point(data = shared_data, aes(x = longitude, y = latitude, color = case_num, size = case_num, 
                                     text = paste("Total Cases: ", case_num, " | Area Name: ", area_name))) +
  coord_sf(
    xlim = c(-79.6303856025883, -79.1180336515019),
    ylim = c(43.5822069220627, 43.8554571861712), expand = FALSE
  ) +
  theme_minimal() +
  labs(title = "Toronto Stolen Bike Case Numbers by Area", x = "Longitude", y = "Latitude", color = "Number of Cases", size = "Total Cases") +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_color_gradientn(colors = c("#603601", "#FFC107"), breaks = color_breaks, labels = rounded_labels) +
  scale_size_continuous(range = c(3, 12))

# Convert ggplot object to interactive plotly object
plotly_map <- ggplotly(ggplot_map, tooltip = "text") %>%
  highlight(
    on = "plotly_hover",
    off = "plotly_doubleclick",
    selected = attrs_selected(line = list(color = "black"))
  )

plotly_map
```
</center>

<div style="color: grey; font-size: 12px; margin-top: 3px; max-width: 800px; margin: 0 auto;">
  Note: The size of each point is proportional to the number of cases reported in each area, highlighting hotspots like downtown and other major regions. By clicking on each point, number of totals cases will be shown for different areas, including Downtown, Etobicoke, York, East York, North York, and Scarborough.
</div> \

<div class="section-heading">
  <span class="section-title-line1">Putting it All Together</span><br>
</div>

<div class="main-body">
  <p>As this article comes to a close, let me reflect on what we’ve uncovered about life as a bike in Toronto. First, we looked at when I’m most at risk. It’s clear that thefts spike in the summer months from <strong style="color: #FFC107;">May</strong> to <strong style="color: #FFC107;">August</strong>, with a noticeable jump on <strong style="color: #FFC107;">Fridays</strong>. These patterns gave us an idea of the timing of bike thefts and helped us understand when extra caution is needed. Next, we dove into recovery rates—or rather, the lack of them. The huge gap between the number of stolen bikes and those recovered highlights just how hard it is for bikes like me to make it back to our owners. It’s a reminder of the need for better tracking and recovery systems.</p>

  <p>We then explored where bike thefts happen, pinpointing hotspots like downtown Toronto and busy transit hubs where the risk is highest. The maps painted a clear picture of these danger zones, giving both cyclists and city planners a guide to where extra security measures are most urgently needed.</p>

  <p>Altogether, these visualizations offer a complete story of Toronto’s bike theft problem. They show the when, where, and why of theft, giving insights that can inspire action. Moving forward, Toronto can do more to protect bikes like me by creating more secure parking, working closely with registration platforms like Project 529, and expanding targeted programs like Project Sentry in high-risk areas. Maybe future studies will take this analysis nationwide, shedding light on bike theft patterns across Canada. With all these efforts combined, cyclists, city officials, and advocacy groups can work together to make Toronto a <strong style="color: #FFC107;">safer</strong> place for bikes like me to ride, rest, and thrive.</p>
</div>

<div class="section-heading">
  <span class="section-title-line1">Conclusion</span><br>
</div>

<div class="main-body">
  <p><strong>This article is a reminder</strong>: as more people turn to biking in Toronto, the risk of <strong style="color: #FFC107;">theft</strong> continues to grow. For bikes like me, the world is full of possibilities—but also risks. Whether you’re a daily commuter or a weekend adventurer, taking extra steps to protect me can go a long way. A sturdy lock isn’t just a piece of metal; it’s a lifeline. Parking in well-lit, high-traffic areas makes thieves think twice. And registering me with platforms like Project 529 ensures I have a better chance of finding my way back home if the worst happens.</p>

  <p>But beyond individual actions, we need a community effort. Cyclists, advocacy groups, and city planners must come together to create a <strong style="color: #FFC107;">safer</strong> Toronto. More secure bike parking, stronger recovery programs, and greater public awareness are the tools we need to keep rolling with confidence. For bikes like me, it’s not just about avoiding theft; it’s about helping Toronto’s cycling community thrive—through every season, every neighborhood, and every ride.</p>

  <p>So as you unlock me and hop on for your next journey, remember this: every precaution you take protects not just me, but the joy of cycling itself. Together, let’s keep Toronto’s streets vibrant, its riders safe, and its bikes where they belong—under the sun, on the move, and by your side.</p>
</div>

<div class="section-heading">
  <span class="section-title-line1"></span><br>
</div>

<div class='citation-section'>
  <h3>References</h3>
  <ul>
    <li>Bike Month. (2024). <i>Bike Month Toronto.</i> <a href='https://www.bikemonth.ca/toronto'>https://www.bikemonth.ca/toronto</a></li>
    <li>City of Toronto. (2019). <i>2019 cycling public opinion survey.</i> <a href='https://www.toronto.ca/wp-content/uploads/2021/04/8f76-2019-Cycling-Public-Option-Survey-City-of-Toronto-Cycling.pdf'>https://www.toronto.ca/wp-content/uploads/2021/04/8f76-2019-Cycling-Public-Option-Survey-City-of-Toronto-Cycling.pdf</a></li>
    <li>City of Toronto. (2024). <i>Bicycle parking.</i> <a href='https://www.toronto.ca/services-payments/streets-parking-transportation/cycling-in-toronto/bicycle-parking/'>https://www.toronto.ca/services-payments/streets-parking-transportation/cycling-in-toronto/bicycle-parking/</a></li>
    <li>CTV News. (2024). Should Toronto tear up its bike lanes to improve traffic flow? Critics say it’s not so simple. <i>CTV News.</i> <a href='https://www.ctvnews.ca/canada/should-toronto-tear-up-its-bike-lanes-to-improve-traffic-flow-critics-say-it-s-not-so-simple-1.7103768'>https://www.ctvnews.ca/canada/should-toronto-tear-up-its-bike-lanes-to-improve-traffic-flow-critics-say-it-s-not-so-simple-1.7103768</a></li>
    <li>Cycle Toronto. (2021). Council talking about ending bike theft week. <i>Cycle Toronto.</i> <a href='https://www.cycleto.ca/council-talking-about-ending-bike-theft-week'>https://www.cycleto.ca/council-talking-about-ending-bike-theft-week</a></li>
    <li>Project 529. (2024). <i>529 garage: Bike registration to reduce theft and help recovery.</i> <a href='https://project529.com/garage'>https://project529.com/garage</a></li>
    <li>Shapiro, A. (2021). Bike thieves are on a roll during the pandemic. Here’s how to protect your ride. <i>NPR.</i> <a href='https://www.npr.org/2021/04/02/980008727/bike-thieves-are-on-a-roll-during-the-pandemic-heres-how-to-protect-your-ride'>https://www.npr.org/2021/04/02/980008727/bike-thieves-are-on-a-roll-during-the-pandemic-heres-how-to-protect-your-ride</a></li>
    <li>Toronto Police Service. (2024a). <i>Bicycle registration.</i> <a href='https://www.tps.ca/services/bicycle-registration/'>https://www.tps.ca/services/bicycle-registration/</a></li>
    <li>Toronto Police Service. (2024b). <i>Bicycle safety & theft prevention.</i> <a href='https://www.tps.ca/crime-prevention/bicycle-safety-and-theft-prevention/'>https://www.tps.ca/crime-prevention/bicycle-safety-and-theft-prevention/</a></li>
    <li>Toronto Police Service. (2024c). <i>Bicycle thefts open data.</i> <a href='https://data.torontopolice.on.ca/datasets/TorontoPS::bicycle-thefts-open-data/about'>https://data.torontopolice.on.ca/datasets/TorontoPS::bicycle-thefts-open-data/about</a></li>
    <li>Toronto Police Service. (2024d). Project Sentry: Keeping Toronto safe. <i>Toronto Police Service.</i> <a href='https://www.tps.ca/media-centre/news-releases/fix-title-project-sentry/'>https://www.tps.ca/media-centre/news-releases/fix-title-project-sentry/</a></li>
  </ul>
</div>